---
title: "621 HW1"
output: pdf_document
date: "2023-01-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2.10 problem 3
## read the data into R
```{r, echo=FALSE}
library(fpp3)
tute1=read.csv("/Users/liangdameinv/Desktop/tute1.csv")
View(tute1)
```
## Convert the data to time series
```{r, echo=FALSE}
mytimeseries <- tute1 %>%
  mutate(Quarter = yearquarter(Quarter)) %>%
  as_tsibble(index = Quarter)
```
## Construct time series plots of each of the three series
```{r, echo=FALSE}
mytimeseries %>%
  pivot_longer(-Quarter) %>%
  ggplot(aes(x = Quarter, y = value, colour = name)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y")
```
## Construct time series plots of each of the three series without 'facet'
```{r, echo=FALSE}
mytimeseries %>%
  pivot_longer(-Quarter) %>%
  ggplot(aes(x = Quarter, y = value, colour = name)) +
  geom_line()
```
## 2.10 problem 4
## Create a tsibble from us_total with year as the index and state as the key.
```{r, echo=FALSE}
library(USgas)
data("us_total")
ts = as_tsibble(us_total, key = state, index = year)
ts
```
## Plot the annual natural gas consumption by state for the New England area (comprising the states of Maine, Vermont, New Hampshire, Massachusetts, Connecticut and Rhode Island).
```{r, echo=FALSE}
library(ggprism)
a=which(us_total$state=="Maine")
b=which(us_total$state=="Vermont")
c=which(us_total$state=="New Hampshire")
d=which(us_total$state=="Massachusetts")
e=which(us_total$state=="Connecticut")
f=which(us_total$state=="Rhode Island")
p1<-ggplot(us_total[a,])+
  geom_line(aes(year, y),size=0.8,color="red")+
  theme_prism(palette = "candy_soft",
              base_fontface = "plain", 
              base_family = "serif", 
              base_size = 16,  
              base_line_size = 0.8, 
              axis_text_angle = 45)
p2<-ggplot(us_total[b,])+
  geom_line(aes(year, y),size=0.8,color="green")+
  theme_prism(palette = "candy_soft",
              base_fontface = "plain", 
              base_family = "serif", 
              base_size = 16,  
              base_line_size = 0.8, 
              axis_text_angle = 45)
p3<-ggplot(us_total[c,])+
  geom_line(aes(year, y),size=0.8,color="pink")+
  theme_prism(palette = "candy_soft",
              base_fontface = "plain", 
              base_family = "serif", 
              base_size = 16,  
              base_line_size = 0.8, 
              axis_text_angle = 45)
p4<-ggplot(us_total[d,])+
  geom_line(aes(year, y),size=0.8,color="yellow")+
  theme_prism(palette = "candy_soft",
              base_fontface = "plain", 
              base_family = "serif", 
              base_size = 16,  
              base_line_size = 0.8, 
              axis_text_angle = 45)
p5<-ggplot(us_total[e,])+
  geom_line(aes(year, y),size=0.8,color="blue")+
  theme_prism(palette = "candy_soft",
              base_fontface = "plain", 
              base_family = "serif", 
              base_size = 16,  
              base_line_size = 0.8, 
              axis_text_angle = 45)
p6<-ggplot(us_total[f,])+
  geom_line(aes(year, y),size=0.8,color="orange")+
  theme_prism(palette = "candy_soft",
              base_fontface = "plain", 
              base_family = "serif", 
              base_size = 16,  
              base_line_size = 0.8, 
              axis_text_angle = 45)
library(cowplot)
plot_grid(p1,p2,p3,p4,p5,p6)

newdata=cbind(us_total[a,],us_total[b,],us_total[c,],us_total[d,],us_total[e,],us_total[f,])
```
## 2.10 problem 7
## Use autoplot(), gg_season() and gg_subseries() to compare the differences between the arrivals from these four countries.
```{r, echo=FALSE}
autoplot(aus_arrivals,Arrivals)+
  labs(y = "arrivals",
       title = "differences between arrivals from four countries")

aus_arrivals%>%
gg_season(Arrivals,labels="both")

aus_arrivals%>%
gg_subseries(Arrivals)
```
## 3.7 problem 1
## Consider the GDP information in global_economy. Plot the GDP per capita for each country over time. Which country has the highest GDP per capita? How has this changed over time?
```{r, echo=FALSE}
global_economy %>%
  autoplot(GDP / Population,show.legend =  FALSE) +
  labs(title= "GDP per capita", y = "$US")

global_economy %>%
  mutate(GDP_per_capita = GDP / Population) %>%
  filter(GDP_per_capita == max(GDP_per_capita, na.rm = TRUE)) %>%
  select(Country, GDP_per_capita)

global_economy %>%
  filter(Country == "Monaco") %>%
  autoplot(GDP/Population) +
  labs(title= "GDP per capita for Monaco", y = "$US")
```
## Monaco has the highest GDP per capital, it has an increasing trend over time.

## 3.7 problem 5
## For the following series, find an appropriate Box-Cox transformation in order to stabilise the variance.    Tobacco from aus_production, Economy class passengers between Melbourne and Sydney from ansett, and Pedestrian counts at Southern Cross Station from pedestrian.
```{r, echo=FALSE}
lambda <- aus_production %>%
  features(Tobacco, features = guerrero) %>%
  pull(lambda_guerrero)
aus_production %>%
  autoplot(box_cox(Tobacco, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Transformed Tobacco use with $\\lambda$ = ",
         round(lambda,2))))

mel_syd <- ansett %>%
  filter(Class == "Economy",
         Airports == "MEL-SYD")
lambda <- mel_syd %>%
  features(Passengers, features = guerrero) %>%
  pull(lambda_guerrero)
mel_syd %>%
  autoplot(box_cox(Passengers, lambda)) +
  labs(y = "", title = latex2exp::TeX(paste0("Transformed Number of Passengers with $\\lambda$ = ",
         round(lambda,2))))

southern_cross <- pedestrian %>%
  filter(Sensor == "Southern Cross Station") 
lambda <- southern_cross %>%
  features(Count, features = guerrero) %>%
  pull(lambda_guerrero)
southern_cross %>%
  autoplot(box_cox(Count, lambda)) +
  labs(y = "", title = latex2exp::TeX(paste0("Transformed Hourly Pedestrian Counts with $\\lambda$ = ",
         round(lambda,2))))
```
## 
```{r, echo=FALSE}
library(astsa)
library(xts)
data=read.csv("/Users/liangdameinv/Desktop/google.csv")
x = read.zoo(data, FUN=as.Date,header=TRUE, sep=',', format = "%Y/%m/%d")
plot(as.xts(x),col=astsa.col(2:3,.7), main='Weekly Google Searches USA')
addLegend("topleft", col=2:3, lty=1, lwd=2, bg=gray(1), bty='o', box.col=gray(1)) 
```
## autoplot() can not only fit ARIMA model, but also it can use the model to forecast the trend.