---
title: "621 hw2"
output: pdf_document
date: "2023-01-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 2

```{r echo=FALSE}
library(astsa)
data=globtemp
plot(data)
lines(ksmooth(time(data),data,"normal",bandwidth = 3),lwd=3,col=4)
par(fig=c(.5,1,.5,1),new=TRUE)
gauss=function(x){1/sqrt(2*pi)*exp(-(x^2)/2)}
x=seq(from=-3,to=3,by=0.001)
plot(x,gauss(x),type="l",ylim=c(-.47,.47),xtat="n",ann=FALSE)
```

```{r echo=FALSE}
library(astsa)
data=globtemp
plot(data)
lines(lowess(data, f=.05), lwd=2, col=4)
lines(lowess(data), lty=2, lwd=2, col=2)
```
## According to the figures, global temperatures are trending upward and are warming at an increasing rate.


## Problem 3

```{r echo=FALSE}
fit = lm(soi~time(soi), na.action=NULL)
par(mfrow=c(1, 1))
plot(resid(fit), type='o', main="Detrended")
lines(fitted(fit), col=2)
lines(lowess(soi, f=.05), lwd=2, col=4)
stats::spec.pgram(soi,
                  fast=FALSE,
                  spans=10,
                  dmean=TRUE,
                  detrend=TRUE,
                  plot=TRUE)
```
## According to the figure, there is a decreasing trend in the sea surface temperature.The less obvious one at the frequency of one cycle every 2 months.


## Problem 4

```{r pressure, echo=FALSE}
library(fpp3)
library(astsa)
data <- aus_production
beer=data.frame(data$Quarter,data$Beer)
beer.ts <- ts(beer, frequency = 4, start = c(1956,1), end = c(2010,2))
plot.ts(beer.ts[,2], main = "Monthly Beer Production in Australia", xlab = "Year", ylab = "ML")
acf(beer.ts[,2])
pacf(beer.ts[,2])
fit = lm(beer.ts[,2]~time(beer.ts[,2]), na.action=NULL)
par(mfrow=c(1, 1))
plot(resid(fit), type='o', main="Detrended")
lines(fitted(fit), col=2)
lines(lowess(beer.ts, f=.05), lwd=2, col=4)
stats::spec.pgram(beer.ts[,2],
                  fast=FALSE,
                  spans=10,
                  dmean=TRUE,
                  detrend=TRUE,
                  plot=TRUE)
```
## According to the figure, the Australian quarterly beer production has an increasing trend.
