---
title: "541hw2"
output: pdf_document
date: "2023-01-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1 a,b

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/CASchools.csv")
A=cov(data[,2:11])
variance=c(rep(0,10))
for(i in 1:10)
  variance[i]=eigen(A)$values[i]/sum(eigen(A)$values)
plot(variance/sum(eigen(A)$values),main="percentage of variance explained by the principle components against the index of the principle component")
```
## According to the plot, I think there is only two factors that can be kept for describing the variation in the dataset.Because the level of explanation of the remaining factors is almost zero.

## Problem 1 c

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/CASchools.csv")
A=cov(data[,2:11])
plot(c(1:10),eigen(A)$vectors[,1])
par(new=TRUE)
points(c(1:10),eigen(A)$vectors[,2],col="red")
par(new=TRUE)
points(c(1:10),eigen(A)$vectors[,3],col="blue")
```
## The plot shows that the first pinciple component explains the data most.

## Problem 1 d

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/CASchools.csv")
A=cov(data[,2:11])
B=c(rep(0,10))
for(i in 1:10)
  B[i]=sd(data[,i+1])
plot(B,eigen(A)$vectors[,1])
```
## According to the plot, we mapped the data to a one-dimensional space.

## Problem 1 e

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/CASchools.csv")
A=cov(data[,2:11])
projection_1 = as.matrix(data[,2:11])%*%eigen(A)$vectors[,1]
projection_2= as.matrix(data[,2:11])%*%eigen(A)$vectors[,2]
projection_3= as.matrix(data[,2:11])%*%eigen(A)$vectors[,3]
plot(projection_2~projection_1,main="PC Plot 1", xlab="Projection onto PC 1",ylab="Projection onto PC 2")
plot(projection_3~projection_1,main="PC Plot 2", xlab="Projection onto PC 1",ylab="Projection onto PC 3")
plot(projection_3~projection_2,main="PC Plot 3", xlab="Projection onto PC 2",ylab="Projection onto PC 3")
```
## According to the three plots, the three principle components are uncorrelated.

## Problem 2 a

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/Hitters.csv")
A=cov(data[,2:14])
variance=c(rep(0,13))
for(i in 1:13)
  variance[i]=eigen(A)$values[i]/sum(eigen(A)$values)
variance
plot(variance/sum(eigen(A)$values),main="percentage of variance explained by the principle components against the index of the principle component")
```
## According to the plot, there is only one principle component that can describe at least 99% of the variance in the data.

## Problem 2 b

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/Hitters.csv")
B=c(rep(0,13))
for(i in 1:13)
  B[i]=sd(data[,i+1])
plot(B,abs(eigen(A)$vectors[,1]))
```
## According to the plot, we mapped the data to a one-dimensional space

## Problem 2 c

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/Hitters.csv")
newdata=scale(data[,2:14])
C=cor(newdata)
variance1=c(rep(0,13))
for(i in 1:13)
  variance1[i]=eigen(C)$values[i]/sum(eigen(C)$values)
variance1
plot(variance1/sum(eigen(C)$values),main="percentage of variance explained by the principle components against the index of the principle component")
```
## There are not principle components can describe at leat 99% of the variation in the data.I would like to keep 4 principle components in order to understand the variation in the data.

## Problem 2 d

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/Hitters.csv")
A=cov(data[,2:14])
D=c(rep(0,13))
for(i in 1:13)
  D[i]=sd(as.matrix(data[,i+1]))
plot(D,abs(eigen(A)$vectors[,1]),xlab = "standard deviation",ylab="first principle component" )
plot(D,abs(eigen(A)$vectors[,2]),xlab = "standard deviation",ylab="second principle component")
plot(D,abs(eigen(A)$vectors[,3]),xlab = "standard deviation",ylab="third principle component")
plot(D,abs(eigen(A)$vectors[,4]),xlab = "standard deviation",ylab="fourth principle component")
```
## The first principle component explain the variation best.

## Problem 2 e 

##The normalized data contains more principal components that can explain the variance, and the variance can be explained more comprehensively with the standardized data.

## Problem 3 a b

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/Artificial1.csv")
A=cov(data[,2:11])
variance=c(rep(0,10))
for(i in 1:10)
  variance[i]=eigen(A)$values[i]/sum(eigen(A)$values)
plot(variance/sum(eigen(A)$values),main="percentage of variance explained by the principle components against the index of the principle component")
projection= as.matrix(data[,2:11])%*%eigen(A)$vectors
projection_1 = as.matrix(data[,2:11])%*%eigen(A)$vectors[,1]
projection_2= as.matrix(data[,2:11])%*%eigen(A)$vectors[,2]
plot(projection_2~projection_1,main="PC Plot 1", xlab="Projection onto PC 1",ylab="Projection onto PC 2")
newdata=scale(data[,2:11])
B=cor(newdata)
variance1=c(rep(0,10))
for(i in 1:10)
  variance1[i]=scale(eigen(B)$values[i]/sum(eigen(B)$values))
projectionx= as.matrix(newdata)%*%scale(eigen(B)$vectors)
projection1 = as.matrix(newdata)%*%scale(eigen(B)$vectors[,1])
projection2= as.matrix(newdata)%*%scale(eigen(B)$vectors[,2])
plot(projection2~projection1,main="PC Plot 1", xlab="Projection onto PC 1",ylab="Projection onto PC 2")
```

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/Artificial2.csv")
A=cov(data[,2:51])
variance=c(rep(0,50))
for(i in 1:50)
  variance[i]=eigen(A)$values[i]/sum(eigen(A)$values)
plot(variance/sum(eigen(A)$values),main="percentage of variance explained by the principle components against the index of the principle component")
projection= as.matrix(data[,2:51])%*%eigen(A)$vectors
projection_1 = as.matrix(data[,2:51])%*%eigen(A)$vectors[,1]
projection_2= as.matrix(data[,2:51])%*%eigen(A)$vectors[,2]
plot(projection_2~projection_1,main="PC Plot 1", xlab="Projection onto PC 1",ylab="Projection onto PC 2")
newdata=scale(data[,2:51])
B=cor(newdata)
variance1=c(rep(0,50))
for(i in 1:50)
  variance1[i]=scale(eigen(B)$values[i]/sum(eigen(B)$values))
projectionx= as.matrix(newdata)%*%(eigen(B)$vectors)
projection1 = as.matrix(newdata)%*%(eigen(B)$vectors[,1])
projection2= as.matrix(newdata)%*%(eigen(B)$vectors[,2])
plot(projection2~projection1,main="PC Plot 1", xlab="Projection onto PC 1",ylab="Projection onto PC 2")
```

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/penguins.csv")
A=cov(data[,2:26])
variance=c(rep(0,25))
for(i in 1:25)
  variance[i]=eigen(A)$values[i]/sum(eigen(A)$values)
plot(variance/sum(eigen(A)$values),main="percentage of variance explained by the principle components against the index of the principle component")
projection= as.matrix(data[,2:26])%*%eigen(A)$vectors
projection_1 = as.matrix(data[,2:26])%*%eigen(A)$vectors[,1]
projection_2= as.matrix(data[,2:26])%*%eigen(A)$vectors[,2]
plot(projection_2~projection_1,main="PC Plot 1", xlab="Projection onto PC 1",ylab="Projection onto PC 2")
newdata=scale(data[,2:26])
B=cor(newdata)
variance1=c(rep(0,25))
for(i in 1:25)
  variance1[i]=scale(eigen(B)$values[i]/sum(eigen(B)$values))
projectionx= as.matrix(newdata)%*%scale(eigen(B)$vectors)
projection1 = as.matrix(newdata)%*%scale(eigen(B)$vectors[,1])
projection2= as.matrix(newdata)%*%scale(eigen(B)$vectors[,2])
plot(projection2~projection1,main="PC Plot 1", xlab="Projection onto PC 1",ylab="Projection onto PC 2")
```
## Problem 3 c
## The effect of principal component analysis for high-dimensional data is not obvious. I prefer to use the plot of the percentage of variance explained by the principle components. Because, it is obvious for us to decide how many principle components should be kept.

## Problem 4 

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/College.csv")
X=as.matrix(data[,2:16])
V=eigen(t(X)%*%(X))$vectors
P=X%*%V
lambda=eigen(t(X)%*%(X))$values
U=eigen(X%*%t(X))$vectors
Q=t(X)%*%U
mu=eigen(X%*%t(X))$values[1:15]
A=svd(X)$u
B=svd(X)$v
gama=svd(X)$d
R=X%*%B
S=t(X)%*%A
gama2=gama^2
```

## Problem 4 a

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/College.csv")
X=as.matrix(data[,2:16])
V=eigen(t(X)%*%(X))$vectors
P=X%*%V
lambda=eigen(t(X)%*%(X))$values
U=eigen(X%*%t(X))$vectors
Q=t(X)%*%U
mu=eigen(X%*%t(X))$values[1:15]
A=svd(X)$u
B=svd(X)$v
gama=svd(X)$d
R=X%*%B
S=t(X)%*%A
gama2=gama^2
all.equal(lambda,mu)
all.equal(lambda,gama2)
```

## Problem 4 b

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/College.csv")
X=as.matrix(data[,2:16])
V=eigen(t(X)%*%(X))$vectors
P=X%*%V
lambda=eigen(t(X)%*%(X))$values
U=eigen(X%*%t(X))$vectors
Q=t(X)%*%U
mu=eigen(X%*%t(X))$values[1:15]
A=svd(X)$u
B=svd(X)$v
gama=svd(X)$d
R=X%*%B
S=t(X)%*%A
gama2=gama^2
a=0
for(i in 1:15)
  {m=cbind(U[,i],R[,i],A[,i],P[,i])
  n=cbind(A[,i],R[,i],P[,i])
  if(qr(m)$rank==qr(n)$rank)
    a=a+1}
a
a=0
for(i in 1:15)
  {m=cbind(U[,i],R[,i],A[,i],P[,i])
  n=cbind(U[,i],P[,i],A[,i])
  if(qr(m)$rank==qr(n)$rank)
    a=a+1}
a
a=0
for(i in 1:15)
  {m=cbind(U[,i],R[,i],A[,i],P[,i])
  n=cbind(U[,i],P[,i],R[,i])
  if(qr(m)$rank==qr(n)$rank)
    a=a+1}
a
a=0
for(i in 1:15)
  i=1
  {m=cbind(U[,i],R[,i],A[,i],P[,i])
  n=cbind(A[,i],P[,i],R[,i])
  if(qr(m)$rank==qr(n)$rank)
    a=a+1}
a
X1=matrix(nrow=3,ncol=15)
for(i in 1:15)
  {n1=cbind(A[,i],P[,i],R[,i])
 sol.svd <- svd(n1)
 U1<-sol.svd$u
 D<-sol.svd$d
 V<-sol.svd$v
 C<-t(U1)%*%U[,i]
 Y<-C/D
 X1[,i]<-V%*%Y}
X1
```
## According to the result, The corresponding columns of the four matrices are linearly correlated, and the corresponding constants can be calculated.

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/College.csv")
X=as.matrix(data[,2:16])
V=eigen(t(X)%*%(X))$vectors
P=X%*%V
lambda=eigen(t(X)%*%(X))$values
U=eigen(X%*%t(X))$vectors
Q=t(X)%*%U
mu=eigen(X%*%t(X))$values[1:15]
A=svd(X)$u
B=svd(X)$v
gama=svd(X)$d
R=X%*%B
S=t(X)%*%A
gama2=gama^2
m=matrix(nrow = 777,ncol=15)
c=c(rep(0,15))
for(i in 1:15)
    {m[,i]=U[,i]/P[,i]
c[i]=unique(m[,i])}
c

m=matrix(nrow = 777,ncol=15)
c=c(rep(0,15))
for(i in 1:15)
    {m[,i]=U[,i]/A[,i]
c[i]=unique(m[,i])}
c

m=matrix(nrow = 777,ncol=15)
c=c(rep(0,15))
for(i in 1:15)
    {m[,i]=U[,i]/R[,i]
c[i]=unique(m[,i])}
c

m=matrix(nrow = 777,ncol=15)
c=c(rep(0,15))
for(i in 1:15)
    {m[,i]=P[,i]/A[,i]
c[i]=unique(m[,i])}
c


m=matrix(nrow = 777,ncol=15)
c=c(rep(0,15))
for(i in 1:15)
    {m[,i]=P[,i]/R[,i]
     c[i]=unique(m[,i])
  }
c


m=matrix(nrow = 777,ncol=15)
c=c(rep(0,15))
for(i in 1:15)
    {m[,i]=A[,i]/R[,i]
c[i]=unique(m[,i])}
c
```
## According to the result, the corresponding columns of U,P,A,R are all constant multiples of each other.

## Problem 4 c

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/College.csv")
X=as.matrix(data[,2:16])
V=eigen(t(X)%*%(X))$vectors
P=X%*%V
lambda=eigen(t(X)%*%(X))$values
U=eigen(X%*%t(X))$vectors
Q=t(X)%*%U
mu=eigen(X%*%t(X))$values[1:15]
A=svd(X)$u
B=svd(X)$v
gama=svd(X)$d
R=X%*%B
S=t(X)%*%A
gama2=gama^2
b=0
for(i in 1:15)
  {m=cbind(V[,i],Q[,i],B[,i],S[,i])
  n=cbind(V[,i],Q[,i],B[,i])
  if(qr(m)$rank==qr(n)$rank)
    b=b+1}
b

b=0
for(i in 1:15)
  {m=cbind(V[,i],Q[,i],B[,i],S[,i])
  n=cbind(V[,i],Q[,i],S[,i])
  if(qr(m)$rank==qr(n)$rank)
    b=b+1}
b

b=0
for(i in 1:15)
  {m=cbind(V[,i],Q[,i],B[,i],S[,i])
  n=cbind(V[,i],S[,i],B[,i])
  if(qr(m)$rank==qr(n)$rank)
    b=b+1}
b

b=0
for(i in 1:15)
  {m=cbind(V[,i],Q[,i],B[,i],S[,i])
  n=cbind(S[,i],Q[,i],B[,i])
  if(qr(m)$rank==qr(n)$rank)
    b=b+1}
b

X2=matrix(nrow=3,ncol=15)
for(i in 1:15)
  {n1=cbind(S[,i],Q[,i],B[,i])
 sol.svd <- svd(n1)
 U1<-sol.svd$u
 D<-sol.svd$d
 V1<-sol.svd$v
 C<-t(U1)%*%V[,i]
 Y<-C/D
 X2[,i]<-V1%*%Y}
X2
```
## According to the result, The corresponding columns of the four matrices are linearly correlated, and the corresponding constants can be calculated.

```{r}
data=read.csv("/Users/liangdameinv/Desktop/541/College.csv")
X=as.matrix(data[,2:16])
V=eigen(t(X)%*%(X))$vectors
P=X%*%V
lambda=eigen(t(X)%*%(X))$values
U=eigen(X%*%t(X))$vectors
Q=t(X)%*%U
mu=eigen(X%*%t(X))$values[1:15]
A=svd(X)$u
B=svd(X)$v
gama=svd(X)$d
R=X%*%B
S=t(X)%*%A
gama2=gama^2

m=matrix(nrow = 15,ncol=777)
c=c(rep(0,15))
for(i in 1:15)
    {m[,i]=V[,i]/Q[,i]
c[i]=unique(m[,i])}
c

m=matrix(nrow = 15,ncol=777)
c=c(rep(0,15))
for(i in 1:15)
    {m[,i]=V[,i]/B[,i]
c[i]=unique(m[,i])}
c

m=matrix(nrow = 15,ncol=777)
c=c(rep(0,15))
for(i in 1:15)
    {m[,i]=V[,i]/S[,i]
c[i]=unique(m[,i])}
c

c=c(rep(0,15))
for(i in 1:15)
    {m[,i]=Q[,i]/B[,i]
c[i]=unique(m[,i])}
c

m=matrix(nrow = 15,ncol=777)
c=c(rep(0,15))
for(i in 1:15)
    {m[,i]=Q[,i]/S[,i]
c[i]=unique(m[,i])}
c

m=matrix(nrow = 15,ncol=777)
c=c(rep(0,15))
for(i in 1:15)
    {m[,i]=B[,i]/S[,i]
c[i]=unique(m[,i])}
c
```
## According to the result, the corresponding columns of U,P,A,R are all constant multiples of each other.