---
title: "Homework #2"
output: pdf_document
fig_height: 5
fig_width: 5
---

#### Instructions
* Produce your assignment as a RMarkdown document
  rendered to pdf (knit to pdf). See instructions in announcements if you have not downloaded a LaTeX distributor.
If you are still having issues, you may knit to HTML as this is the first assignment.
However, future assignments that are not knitted to pdf will be penalized.
  
* Also submit your Rmd file (it will not be graded but we want it for
  reference purposes).
* Show all the code (use `echo = TRUE` as option in R chunks) as
well as the results.
* 80 total points (20 points per exercise).
* See Syllabus for HW policies.

#### Exercise 1:
* Make sure you have the library ggplot2 installed.
If not, install it by running the uncommented code below:
```{r, echo = TRUE}
## install.packages("ggplot2")
```

* Run the code below to access and prepare the data:
```{r, echo = TRUE}
## Load the library
library(ggplot2)
## Activate the midwest data.frame provided by ggplot2
data(midwest)
## New versions of ggplot provide a tibble and have all character variables, while
## older were data frames and had factors. Transform state to factor
## to show how most data.frames treat character variables.
midwest <- as.data.frame(midwest)
midwest$state <- factor(midwest$state)
```

1. Inspect the structure of the midwest data.frame. Note the kind of data provided.
2. Run the summary function to learn more about the variables.
3. Get a subset of the data.frame including all counties that are not in state `MI`. No need to show the dataset here.
4. Determine if the state variable (that is a factor) in your subset has dropped the now-removed state `MI` hint:use the table function.
5. Devise a strategy to assure that the above factor has dropped the level that has no elements. I.e. if levels() function is used, `MI` should not show up.
6. Further subset the data making sure that only counties with an `area` at least 0.06, `perchsd` less than or equal to 80, and EITHER `percchildbelowpovert` greater than 10 OR `percadultpoverty` greater than 10 are included. Do not show the dataset. Instead, print the dimensions.
7. Determine how many counties per state satisfy these constraints. Only show states with at least one county.
8. Devise a strategy to only show the state(s) with most counties (Note: your solution should also contemplate the possibility of a tie for the first place and should not include state names).

```{r}
## Begin solution
str(midwest)

summary(midwest)

which(midwest$state!="MI")
frame1=midwest[which(midwest$state!="MI"),]

table(frame1$state!="MI")

f=factor(frame1$state)
fa=levels(f)
fa

frame2=frame1[which(frame1$area>=0.06&frame1$perchsd<=80&(frame1$percchildbelowpovert>=10|frame1$percadultpoverty>=10)),]
dim(frame2)

data2=frame2[,2:3]
ct=c(rep(0,4))
for(i in 1:4)
  ct[i]=sum(data2$state==fa[i])
ct
for(i in 1:4)
  if(ct[i]>=1)
    print(fa[i])

t=function(x){
  k=1
  m=x[1]
  for(i in 2:4)
    if(x[i]>m)
      {m=x[i]
      k=i}
k
}
t(ct)
## End solution
```


#### Exercise 2
\
Print the results for each problem.

1. Create a list `L` that has a vector c(1,2,3) as the first element and a vector c(4,5,6) as its second element.
2. Use subsetting to replace the 6 with a 7.
3. Merge another vector c(6,8,9) to `L`.
4. Devise a strategy to take the sum of the elements without explicitly subsetting them (i.e. do not use [[1]], [[2]], or [[3]]). hint: convert the list to a vector.

```{r}
## Begin solution
c1=c(1,2,3)
c2=c(4,5,6)
L=list(c1,c2)

l=subset(L[[2]],L[[2]]>3)
l
l1=replace(l,l>5,7)
l1
L=list(c1,l1)

c3=c(6,8,9)
L=list(c1,l1,c3)

l1=L[[1]]
l2=L[[2]]
l3=L[[3]]
r=rbind(l1,l2,l3)
r1=as.matrix(r)
r1
a=c(rep(0,3))
for(i in 1:3)
  a[i]=sum(r1[i,])
a
## End solution
```


#### Exercise 3
\
Print the results for each problem.

1. Create a 2x12 (2 rows 12 cols) matrix `M` using the built in vectors `month.abb` and `month.name`.
2. Devise a strategy to merge the vectors in 1) with alternating indices. I.e. the vector should print "Jan" "January" "Feb" "February" etc. Hint: Use the matrix you created.
3. Name the columns of `M` the corresponding numeric number (1 for Jan, 2 for Feb, etc.) in a single line of code.

```{r}
## Begin solution
month.name
m=rbind(month.abb,month.name)
M=matrix(data=m,nrow = 2,ncol=12)
M

m1=c(rep(0,24))
y=function(x){
  n=1
  for(i in 1:12)
    {for(j in 1:2)
      {m1[n]=x[j,i]
      n=n+1}}
m1
}
y(M)

M[1,]=c(1:12)
M
## End solution
```

#### Exercise 4
\
`iris` is a built in dataset that contains information about 3 different species of iris (Use ?iris for more information).

1. Print the species from highest to lowest average `Petal.Length`. Do this without ever having to manually input
a sepal name (I.e. do not use "virginica", etc. in your code).
2. The iris with the highest sum of `Sepal.Width` and `Petal.Length` belongs to which species? You may find it useful to use `which.max` here.

```{r}
## Begin Solution
data3=iris
data=aggregate(data3$Petal.Length,by=list(data3$Species),FUN =mean )
data1=data[order(-data$x),]
data1

data3[which.max(data3$Sepal.Width+data3$Petal.Length),5]
## End Solution
```


