---
title: "Homework #3"
output: pdf_document
fig_height: 5
fig_width: 5
---

#### Instructions
* Produce your assignment as a RMarkdown document
  rendered to pdf (knit to pdf).
* Also submit your Rmd file (it will not be graded but we want it for
  reference purposes).
* Show all the code (use `echo = TRUE` as option in R chunks) as
well as the results.
* Make sure your code can be generalized, when applicable.
* 100 total points.
* See Syllabus for HW policies.

#### Exercise 1 (20 points):
* Make your own quantile-quantile plot. You will have
  to compare the quantiles of the theoretical
  distribution (in this case a Normal distribution with mean
  equal the sample mean of the data and standard deviation equal the sample
  standard deviation of the data), with your sorted data.
  
* Hint 1: use the `qnorm` function,
  consulting its help for needed parameters. Of course, you cannot use `qqnorm`
  in the code you produce, but you can use it to compare with your results.
  
* Hint 2: to find the theoretical quantiles, you will have to
        create a vector of probabilities with the same number
        of elements as your data, excluding the extremes
        0 and 1.
* Test it on the following data:

```{r, echo = TRUE}
set.seed(123)
r <- rnorm(1000)
```


```{r, fig.align = 'center'}

```

#### Exercise 2 (20 points):
* Adapt the quantile-quantile procedure
  developed above to compare to a
  Gamma distribution with the following data. 

```{r, echo = TRUE}
set.seed(123)
r <- rgamma(1000, shape = 1)
```
  
```{r, fig.align = 'center'}

```

#### Exercise 3 (20 points):
* Using `faithful` dataset, do a stem and leaf
 plot (`stem`), a histogram to which you will add a `rug`,
 and a boxplot, using in all cases the variable `eruptions`.
 Based on the data alone, does it appear normally distributed?
 Check using a test if it can be concluded or not that
 the eruptions are normally distributed.

```{r, fig.align = 'center'}

```

#### Exercise 4 (40 points):

```{r, message=FALSE, warning=FALSE}
library(UsingR)
data(father.son)
lmfs <- lm(sheight ~ fheight, data = father.son)
```

1. Create a scatterplot of fathers heights (on x axis) and sons heights 
(on y axis).

2. Extract the coefficients of the regression line
from lmfs, and add the regression line IN RED
to the scatterplot (do not use abline(lmfs)).

3. Plot the father heights on x and the residuals in y. 
Add a horizontal line at 0. Why is this step important in a regression?

4. Extract the tstatistic from fheight from the summary
and perform the calculation to obtain
the p-value
i.e., the line that says
fheight      0.51409    0.02705   19.01   <2e-16 ***
Hint: use the function "pt" that gives you the 
cumulative probability of an t distribution. Recall
what the degrees of freedom for a t distribution is
and where it appears in the summary (You can input the df
manually). Additionally, think about how the p-value is actually
calculated.

5. Extract the f statistic from the summary
and perform the calculation to obtain
the p-value of the anova of the regression
i.e., the line that says
F-statistic: 361.2 on 1 and 1076 DF,  p-value: < 2.2e-16
Hint: use the function "pf" that gives you the 
cumulative probability of an F distribution.

```{r}

```
