---
title: "Pledged Individual Midterm"
author: "Yuwen Liang"
date: "2022-10-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1 Matrix and Vector Operations.


```{r}
X=matrix(data=c(1,1,1,5,4,3,2,3,2),ncol=3)
y=c(1,2,3)
```

* 1. Compute $X^2$.

```{r}
X_2=X%*%X
X_2
```
* 2. Compute $X^T$.
```{r}
X_T=t(X)
X_T
```
 * 3. Compute $X^{-1}$.
```{r}
X_1=solve(X)
X_1
```
`

* 4. Compute the equation.
```{r}
X1=solve(t(X)%*%X)%*%t(X)%*%y
X1
```
 ``

* 5. Compute the equation.
```{r}
I=diag(1,3)
I
X2=solve(t(X)%*%X+I)%*%t(X)%*%y
X2
```
  * 6. Take the 10th power of X using eigen-decomposition (eigen() function in R). 
```{r}
X=matrix(data=c(1,1,1,5,4,3,2,3,2),ncol=3)
X.eigen=eigen(X)
c=X.eigen$values
c=c^10
y=X.eigen$vectors%*%diag(c)%*%solve(X.eigen$vectors)
y
```
  
  * 7. For this problem, use sapply() to construct this matrix difference operator B. Then  compute the row-wise difference.  
```{r}
X=matrix(data=c(1,1,1,5,4,3,2,3,2),ncol=3)
B=sapply(1:3,function(i)diff(diag(3)[,i]))
B=as.matrix(B)
B
y=B%*%X
y
```
    
  * 8. Recover the row of the largest entry in X using the mod function. mod is %% in R.  
```{r}
X=matrix(data=c(1,1,1,5,4,3,2,3,2),ncol=3)
y=X%%max(X)
which(y==0,arr.ind=T)[1]
```

* 9. Recover the column of the largest entry in X using the mod function.  
```{r}
X=matrix(data=c(1,1,1,5,4,3,2,3,2),ncol=3)
y=X%%max(X)
which(y==0,arr.ind=T)[2]
```

## Preparation for program 2 and program 3.


```{r}
library(knitr)
library(tidyverse)
library(ggplot2)
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
theme1 <- theme_bw() +
  theme(axis.text = element_text(size = 8, colour = "#6b3447"),
        axis.title = element_text(size = 10, colour = "#2f2f63"),
        legend.title = element_text(size = 8, colour = "#2f2f63"),
        legend.text = element_text(size = 8, colour = "#6b3447"),
        title = element_text(size = 12, colour = "#2f2f63"),
        axis.ticks = element_line(colour = "#6b3447"),
        plot.caption = element_text(size = 8, colour = "#2f2f63"),
        plot.subtitle = element_text(size = 10, colour = "#2f2f63"))
```

## Problem 2 use ggplot to get a plot.

```{r}
data_1116=read.csv('/Users/liangdameinv/Desktop/605/midterm---2.csv')
product=unique(data_1116$product)
ndata=aggregate(data_1116$product,by=list(data_1116$Year,data_1116$product,
                                          data_1116$submitted_via),length)
ndata1=data.frame(Year=ndata$Group.1,product=ndata$Group.2,submitted_via=ndata$Group.3,
                  count=ndata$x)
```

### 1-3 answer.

```{r,echo=TRUE,fig.align='center',fig.width=11,fig.height=15}

p=ggplot(ndata1,aes(x=product,y=count,fill=submitted_via))+
  geom_bar(stat='identity',position='stack')+
  facet_wrap(.~Year,ncol=3)+theme1+
  theme(strip.text.x=element_text(size=15))

```


### 4 answer.


```{r,echo=TRUE,fig.align='center',fig.width=11,fig.height=15}

p1=p+theme(axis.text.x=element_text(angle=90,hjust=1,size=13),
           axis.text.y=element_text(size=13),
           axis.title.x = element_text(size=17),
           axis.title.y = element_text(size=17))

```

### 5-7 answer.

```{r,echo=TRUE,fig.align='center',fig.width=11,fig.height=15}

p1+labs(x='Product',y='Number of Complains')+
  ggtitle('Number of Complaints by Product, Submission Method, and Year')+
  theme(plot.title=element_text(size=17))+
  scale_fill_manual(values=cbPalette)

```


## Problem 3 Use ggplot to create a bar graph for the number of complaints. 

###  1 answer.

```{r}
data=read.csv("/Users/liangdameinv/Desktop/605/midterm---3.csv")
newdata=subset(data,data$state=="CA"|data$state=="FL"|data$state=="MO"|data$state=="TX")
new=data.frame(state=newdata$state,submitted_via=newdata$submitted_via)
new1=aggregate(new$state,by=list(new$submitted_via,new$state),length)
new2=data.frame(state=new1$Group.2,submitted_via=new1$Group.1,count=new1$x)
```

###  2-3 answer.

```{r}
p=ggplot(new2,aes(x=submitted_via,y=count,fill=submitted_via))
p1=p + geom_col() +
  guides(fill = "none") +theme1+
  facet_wrap(~state,ncol=4, scales="free_y")+
  theme(strip.text.x=element_text(size=15))

```


###  4 answer.

```{r,echo=TRUE,fig.align='center',fig.width=11,fig.height=9}
p1=p1+theme(axis.text.x=element_text(angle=35,hjust=1,size=13),
         axis.text.y = element_text(size=13),
         axis.title.x = element_text(size=17),
         axis.title.y = element_text(size=17),
         axis.title = element_text(size=17))

```

### Problem 3 6-8 answer.


```{r,echo=TRUE,fig.align='center',fig.width=11,fig.height=9}
p1+labs(x='Submission Method',y='Number of Complains')+
  ggtitle('Number of Complaints by Submission Method and State')+
  theme(plot.title = element_text(size=16.5)
        )+
  guides(fill = "none")+
  scale_fill_manual(values=cbPalette)

```


## Problem 4  plot the densities of the mpg of the different kinds of cyl.  
```{r,echo=TRUE,fig.align='center',fig.width=5,fig.height=5.3}
##begin
mdata=data.frame(mpg=mtcars$mpg,cylinders=mtcars$cyl)
mdata$cylinders=factor(mdata$cylinders)
m=aggregate(mtcars$mpg,by=list(mtcars$cyl),mean)
m1=data.frame(cyl=m$Group.1,x=m$x)
m1$cyl=factor(m1$cyl)
p=ggplot(data=mdata, aes(x=mpg,color=cylinders)) +
  geom_density(adjust=1.5, alpha=1,lwd=1,linetype = 1)+
  geom_vline(data=m1,aes(xintercept=x,color=cyl),linetype='dashed',lwd=1.5)+
  ylim(0.0,0.25)+
  labs(x='Miles per Gallon')+
  ggtitle("MPG Density Curves by Number of Cylinders")
p+theme_bw() +
  theme(legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.key.height=unit(1.2,'cm'),
        legend.key.width=unit(1.2,'cm'),
        plot.title = element_text(family = "serif",
                                  size=14,
                                  face='bold'))
##end
```

## Problem 5 hypothesis testing.
### Define the function of plot_basic
  
  

```{r}
plot_basic=function(x,dvalues,titlename,label){
    plot(x,dvalues, 
       type = "l",
       main = titlename,
       ylim=c(0,0.45),
       xlim=c(0,8),
       xlab= "x",
       ylab="Density",
       lwd=4) 
  abline(h=0,lwd=2)
  abline(v=mu0,lwd=1)
  legend('topleft',label,title.cex = 0.5,x.intersp = 0.2)
}
```


### Define the function of plot_crit
  
 
```{r}
plot_crit=function(l,r){
  stdev0=stdev/sqrt(n)
  x=seq(-2,10,0.1)
  dvalues=dnorm(x,mu0,stdev0)
  x.p=c(x[l],x[l:r],x[r])
  y.p=c(0,dvalues[l:r],0)
  polygon(x.p,y.p,col="red",angle=-45,density = 10,lwd=2)
  i3=which.min(abs(x-sample.mean))
  segments(sample.mean,0,sample.mean,dvalues[i3],col="orange",lwd=4)
}
```
  

### Define the function of plot_pvalue
 

```{r}
plot_pvalue=function(l,r){
    stdev0=stdev/sqrt(n)
    x=seq(-2,10,0.1)
    dvalues=dnorm(x,mu0,stdev0)
    x.p=c(x[l],x[l:r],x[r])
    y.p=c(0,dvalues[l:r],0)
    polygon(x.p,y.p,col="orange",angle=45,density=10,lwd=2)
}

```
  

### Define the function of plot_blue

  
```{r}
plot_blue=function(l,r){
    x=seq(-2,10,0.1)
    x2=seq(-2,10,0.1)
    stdev0=stdev/sqrt(n)
    dvalues=dnorm(x,mu0,stdev0)
    stdev0=stdev/sqrt(n)
    dvalues1=dnorm(x2,mu1,stdev0)
    x.p=c(x2[l],x2[l:r],x2[r])
    y.p=c(0,dvalues1[l:r],0)
    polygon(x.p,y.p,col="blue")
    lines(x,dvalues,lwd=4)
    abline(v=mu0,lwd=1)
}
```
  

### Define the function of plot_alt

  
```{r}
plot_alt=function(){
    x2=seq(-2,10,0.1)
    x=seq(-2,10,0.1)
    stdev0=stdev/sqrt(n)
    dvalues1=dnorm(x2,mu1,stdev0)
    lines(x2,dvalues1)#density of mu1
    abline(v=mu1,lwd=1)
    i1=which.min(abs(x-qnorm(1-signif.level,mu0,stdev0)))
    i4=which.min(x2)
    i5=which.min(abs(x2-x[i1]))
    plot_blue(i4,i5)#blue region
}

```


### Define the function of plot_beta
 

```{r}
plot_beta=function(){
  x2=seq(-2,10,0.1)
  x=seq(-2,10,0.1)
  stdev0=stdev/sqrt(n)
  dvalues=dnorm(x,mu0,stdev0)
  dvalues1=dnorm(x2,mu1,stdev0)
  i1=which.min(abs(x-qnorm(1-signif.level,mu0,stdev0)))
  i5=which.min(abs(x2-x[i1]))
  plot(x2,pnorm(x2[i5],x2,stdev0),type="l",ylim=c(-0.1,1.1),xlim=c(0,8)
         ,col="blue",lwd=3,xlab="mu1",ylab="Error type II")
  i6=round(pnorm(x2[i5],mu1,stdev0),2)
  text(5.8,pnorm(x2[i5],mu1,stdev0),i6)
  abline(v=mu1)
}

```
  

### Define the function of plot_power
  

```{r}
plot_green=function(l,r){
    stdev0=stdev/sqrt(n)
    x2=seq(-2,10,0.1)
    dvalues1=dnorm(x2,mu1,stdev0)
    x.p=c(x2[l],x2[l:r],x2[r])
    y.p=c(0,dvalues1[l:r],0)
    polygon(x.p,y.p,col="dark green")
    lines(x2,dvalues1)
    abline(v=mu1,lwd=1)
}
```
  
  
```{r}
plot_power=function(){
    x2=seq(-2,10,0.1)
    x=seq(-2,10,0.1)
    stdev0=stdev/sqrt(n)
    dvalues1=dnorm(x2,mu1,stdev0)
    i1=which.min(abs(x-qnorm(1-signif.level,mu0,stdev0)))
    i2=which.max(x)
    i3=which.min(abs(x-sample.mean))
    i4=which.min(x2)
    i5=which.min(abs(x2-x[i1]))
    i6=round(pnorm(x2[i5],mu1,stdev0),2)
    i7=which.max(x2)
    plot_green(i5,i7)#green region
    plot_crit(i1,i2)#red region
    plot_pvalue(i3,i2)#orange region
    plot_blue(i4,i5)#blue region
}

```



### Define the function of plot_error of one side
 

```{r}
plot_error_1=function(){
    x2=seq(-2,10,0.1)
    x=seq(-2,10,0.1)
    stdev0=stdev/sqrt(n)
    i1=which.min(abs(x-qnorm(1-signif.level,mu0,stdev0)))
    i5=which.min(abs(x2-x[i1]))
    plot(x2,pnorm(x2[i5],x2,stdev0),type="l",ylim=c(-0.1,1.1),xlim=c(0,8)
     ,col="blue",lwd=3,xlab="mu1",ylab="Error type II")
    i6=round(pnorm(x2[i5],mu1,stdev0),2)
    text(5.9,pnorm(x2[i5],mu1,stdev0),i6)
    abline(v=mu1)
    lines(x2,1-pnorm(x2[i5],x2,stdev0),col="dark green",lwd=3)
    i8=round(1-pnorm(x2[i5],mu1,stdev0),2)
    text(6.6,1-pnorm(x2[i5],mu1,stdev0),i8)
}

```
  

### Define two side error type.
   
  
  
```{r}
twosided_error=function(){
x2<- seq(-2,10,0.1)
x<-seq(-2,10,0.1)
stdev=stdev/sqrt(n)
dvalues <- dnorm(x,mu0,stdev)
i1=which.min(abs(x-qnorm(1-signif.level,mu0,stdev)))
i10=which.min(abs(x-qnorm(signif.level,mu0,stdev)))
i5=which.min(abs(x2-x[i1]))
i12=which.min(abs(x2-x[i10]))
i6=round(pnorm(x2[i5],6.2,stdev)-pnorm(x2[i12],6.2,stdev),2)
i8=round(1-pnorm(x2[i5],6.2,stdev)-pnorm(x2[i12],6.2,stdev),2)
plot(x2,pnorm(x2[i5],x2,stdev)-pnorm(x2[i12],x2,stdev),type="l",ylim=c(-0.1,1.1),xlim=c(0,8)
     ,col="blue",lwd=3,xlab="mu1",ylab="Error type II")
text(5.9,pnorm(x2[i5],6.2,stdev)-pnorm(x2[i12],6.2,stdev),i6)
abline(v=6.2)
lines(x2,1-(pnorm(x2[i5],x2,stdev)-pnorm(x2[i12],x2,stdev)),col="dark green",lwd=3)
text(6.6,1-pnorm(x2[i5],6.2,stdev)-pnorm(x2[i12],6.2,stdev),i8)
}


```


### Define the function add_left.
 

```{r}
add_left=function(){
  x<- seq(-2,10,0.1)
  stdev0=stdev/sqrt(n)
  dvalues <- dnorm(x,mu0,stdev0)
  i1<- which.min(abs(x-qnorm(0.975,4,stdev0)))
  i9=which.min(x)
  i10=which.min(abs(x-qnorm(0.025,4,stdev0)))
  m=2*mu0-sample.mean
  i11=which.min(abs(x-m))
  x5.p=c(x[i9],x[i9:i10],x[i10])
  y5.p=c(0,dvalues[i9:i10],0)
  segments(x[i10],0,x[i10],dvalues[i10],col="red",lwd=2)
  polygon(x5.p,y5.p,col="red",angle=-45,density=10,lwd=2)
  x6.p=c(x[i9],x[i9:i11],x[i11])
  y6.p=c(0,dvalues[i9:i11],0)
  segments(m,0,m,dvalues[i11],col="orange",lwd=2)
  polygon(x6.p,y6.p,col="orange",angle=45,density=10,lwd=2)
}


```

  

## Define the function of plot hyp-testing
  

```{r}
hyp.testing=function(mu0,stdev,signif.level,sample.mean,n,show_crit='n',
                     show_pvalue='n',show_alt='n',mu1,show_beta='n',
                     show_power='n',two_sided='n'){
  x<- seq(-2,10,0.1)
  stdev0=stdev/sqrt(n)
  dvalues = dnorm(x,mu0,stdev0)
  z=(sample.mean-mu0)/stdev0
  if(two_sided=='n'){
    crit=round(qnorm(1-signif.level,mu0,stdev0),2)
    pvalue=round(1-pnorm(z),6)
    titlename =paste('H0: mu = ',mu0,' vs H1: mu >', mu0)
    lab1=paste("critical region sup =",crit)
    lab2=paste("sample statistics =",sample.mean)
    lab3=paste("p-value =",pvalue)
    label=c(lab1,lab2,lab3)
  }
  else{
    i1=which.min(abs(x-qnorm(1-signif.level,mu0,stdev0)))
    i10=which.min(abs(x-qnorm(signif.level,mu0,stdev0)))
    pvalue=round(2*(1-pnorm(z)),6)
    titlename =paste('H0: mu = ',mu0,' vs H1: mu !=', mu0)
    lab1=paste("critical region sup =",x[i1])
    lab2=paste("critical region inf =",x[i10])
    lab3=paste("sample statistics =",sample.mean)
    lab4=paste("p-value =",pvalue)
    label=c(lab1,lab2,lab3,lab4)
  }
  plot_basic(x,dvalues,titlename,label)
  if(show_crit=='y'){
    l=which.min(abs(x-qnorm(1-signif.level,mu0,stdev0)))
    r=which.max(x)
    plot_crit(l,r)
  }
  if(show_pvalue=='y'){
    l=which.min(abs(x-sample.mean))
    r=which.max(x)
    plot_pvalue(l,r)
  }
  if(show_alt=='y'){
    plot_alt()
    if(show_beta=='y')
    {plot_beta()}
  }
  if(show_power=='y'){
    plot_power()
    plot_error_1()
  }
  if(two_sided=='y'){
    plot_power()
    add_left()
    twosided_error()
  }

}
```


```{r}
mu0 <- 4 
stdev <- 3
signif.level <- 0.05
sample.mean <- 6.07 
n <- 10 
mu1<-6.2
```

## 1. plot crit

```{r}
hyp.testing(mu0,stdev,signif.level=0.05,sample.mean,n,show_crit='y')

```
  
## 2. plot pvalue

```{r}
hyp.testing(mu0,stdev,signif.level=0.05,sample.mean,n,show_crit='y',show_pvalue = 'y')
```

## 3. Add the error of type

```{r}
hyp.testing(mu0,stdev,signif.level=0.05,sample.mean,n,show_crit='y',show_pvalue = 'y',
            show_alt = 'y',mu1,show_beta='y')
```

## 4. Add the power

```{r}
hyp.testing(mu0,stdev,signif.level=0.05,sample.mean,n,show_crit='y',show_pvalue = 'y',
            show_alt = 'y',mu1,show_power = 'y')
```



## 5.Create the two-sided version

```{r}
signif.level=0.025
hyp.testing(mu0,stdev,signif.level=0.025,sample.mean,n,mu1,two_sided = 'y')
```

